\propertitle

### 2019-11-19 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-18T22:37:13+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
log-ratio of the likelihoods of the hypotheses $M''$ and $M'$ in view of
the data $\yf$:
\begin{equation}
  \label{eq:bayes_factor}
  \ydi(M'',M') =
  \log\bigl[\pf(\yf \| M'')/
  \pf(\yf \| M')\bigr].
\end{equation}
This is called their \emph{relative weight of evidence}, the logarithm of
their \emph{relative Bayes factor}
\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells us
how much more probable the data $\yf$ are conditional on $M''$, than
conditional on $M'$. We can also combine this measure with pre-data
probabilities for the two hypotheses to obtain the ratio of their
probabilities conditional on the data.

@@

The application at the full-network level therefore leads to different
conclusions about the relative importance of the moments from the
application at the sample level. It is also possible to use the
measure~\eqref{eq:measure_suffic} and its probabilistic
meaning~\eqref{eq:bayes_factor} to compare the probability of the data
$\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
the full-network level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
5-moment sufficiency at the sample level, $N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi(M_{5, N}, M_{5, n}) = 35\;\nat = 15\;\hart,
\end{equation}
that is, the data are 15 orders of magnitude more likely under the first
hypothesis than under the second.

@@

The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
our method is based, and which asserts the equality of the factorial
moments of the full network and those of the sample. This formula is only
valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
of the full network $\pxx(\yA)$ are related by a hypergeometric
distribution:
\begin{equation}
  \label{eq:hypergeometric_relation}
    \pxx(\ya)=
  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
\;\pxxx(\yA),
\end{equation}
characteristic of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, $N$ is the size of the pool of all neurons that could just
as well have been recorded by our measuring procedure. \enquote{Just as
  well} indicates equal degrees of belief. Whether the equality of such
beliefs is justified, or for which sampling procedures it could be
justified, is the fundamental, deep question of sampling theory, which we
cannot discuss here of course \citep[see \eg\ the discussions and
references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
such as: the shape and dimensions of the recording probe; the inaccuracy in
its insertion, leading for example to slightly different insertion points
or angles; the density of neurons around the point of insertion of the
probe.

Its order of magnitude is enough for making qualitative inferences. As
discussed in the next section, even if the exact number were known, the
maximum-entropy distribution ought to be interpreted qualitatively.

It would be possible to derive a distribution $\pxx(A)$ for the full
network based on an unknown $N$, by  expressing our belief distribution
$\pxx(N)$ about $N$ and using this to marginalize $N$ out:
\begin{equation}
  \label{eq:marginalize_N}
  \pxx(A \| \text{$N$ \small unknown}) = \sum_{N} \px(\yA \| N)\;\pxx(N).
\end{equation}
But this would be overkill, owing to the qualitative character of the
maximum-entropy distribution constructed with exact $N$.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-19T08:53:02+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
log-ratio of the probabilities of the data $\yf$ conditional on the
hypotheses $M''$ and $M'$:
\begin{equation}
  \label{eq:bayes_factor}
  \ydi(M'',M') =
  \log\bigl[\pf(\yf \| M'')/
  \pf(\yf \| M')\bigr].
\end{equation}
This is called their \emph{relative weight of evidence}, the logarithm of
their \emph{relative Bayes factor}
\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells us
how much more probable the data $\yf$ are conditional on $M''$, than
conditional on $M'$. We can also combine this measure with pre-data
probabilities for the two hypotheses to obtain the ratio of their
probabilities conditional on the data.

@@

The application at the full-network level therefore leads to different
conclusions about the relative importance of the moments from the
application at the sample level. It is also possible to use the
measure~\eqref{eq:measure_suffic} and its probabilistic
meaning~\eqref{eq:bayes_factor} to compare the probability of the data
$\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
the full-network level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
5-moment sufficiency at the sample level, $N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi(M_{5, N}, M_{5, n}) = 35\;\nat = 15\;\hart,
\end{equation}
that is, the data are 15 orders of magnitude more probable under the first
hypothesis than under the second.

@@

The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
our method is based, and which asserts the equality of the factorial
moments of the full network and those of the sample. This formula is only
valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
of the full network $\pxx(\yA)$ are related by a hypergeometric
distribution:
\begin{equation}
  \label{eq:hypergeometric_relation}
    \pxx(\ya)=
  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
\;\pxxx(\yA),
\end{equation}
characteristic of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, $N$ is the size of the pool of all neurons that could just
as well have been recorded by our measuring procedure. \enquote{Just as
  well} indicates equal degrees of belief. Whether the equality of such
beliefs is justified, or for which sampling procedures it could be
justified, is the fundamental, deep question of sampling theory, which we
cannot discuss here of course \citep[see \eg\ the discussions, reviews, and
references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
such as: the shape, dimensions, and technical specifications of the
recording probe; the inaccuracy in the insertion of the probe, leading for
example to slightly different insertion points or angles; the density of
neurons around the probe; and similar. $N$ can only be assessed case by
case. The plots of \fig~\ref{fig:plots_5mom} suggest that its order of
might be enough for qualitative inferences. As discussed in the next
section, even if the exact number $N$ were known, the maximum-entropy
distribution ought to be interpreted qualitatively or semi-quantitatively.

If our uncertainty about $N$ spans several orders of magnitude, we can of
course assign a probability to the possible values of $N$ based on our
background information and the frequency data $\yf$. The procedure is
similar to that of \sect~\ref{sec:marginalization} and appendix***. The
probability for the data $\yf$ conditional on a set of constraints $M$ and
$N$ is the exponential of the relative entropy~\eqref{eq:relentropy_T}:
\begin{equation}
  \label{eq:prob_data_from_N}
  \pf(\yf \| M, N) \approx \exp\bigl\{
  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}
If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
\begin{equation}
  \label{eq:prob_N_from_data}
  \pf(N \| M, \yf) \propto p(N)\;\exp\bigl\{
  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}

Here is a simple illustrative example with our first data set. Suppose our
uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
to $N=20\,000$. Divide this range roughly into thirds of order of
magnitude, considering the values
$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming five
moments $M$ are sufficient,
formula~\eqref{eq:prob_data_from_N} gives
\begin{equation}
  \label{eq:data1_f_from_N}
  \begin{gathered}
  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
  \pf(\yf \| N=2\,000, M) =  0.00704,\\
  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
  \pf(\yf \| N=10\,000, M) =  0.0150,\\
  \pf(\yf \| N=20\,000, M) =  0.0135.
\end{gathered}
\end{equation}
Our uncertainty is about a scale factor, so it can be represented by equal
pre-data probabilities $1/5$ about these partial orders of magnitude. Thus
from~\eqref{eq:data1_f_from_N} we find 
\begin{equation}
  \label{eq:data1_N_from_f}
  \begin{gathered}
  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
  \pf(N=2\,000 \| \yf, M) =  0.140,\\
  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
  \pf(N=10\,000 \| \yf, M) =  0.298,\\
  \pf(N=20\,000 \| \yf, M) =  0.267,
\end{gathered}
\end{equation}
which gives a slightly higher probability to $N=10\,000$.
